---
title       : NoSQL 
subtitle    : 
author      : Ahsan Ijaz
job         : Ebryx
# logo        : ebryx-logo.png
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow   # 
# url:
  # lib: ../../libraries
  # assets: /home/ahsan/Personal/Workshop Machine learning/Workshop1 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
embed : TRUE
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
library(knitr)
# make this an external chunk that can be included in any file
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})

knit_hooks$set(plot = knitr:::hook_plot_html)
```
## NoSQL overview

```{r xtable1, results="asis", echo = FALSE, fig.height=52, fig.width=7}
databases = c("RDBMS","Memcached", "MapReduce", "CouchDB", "BigTable (HBase)", "MongoDB", "Dynamo","Pig","HIVE","Cassendra","Voldemort","Riak","Dremel","Megastore","Tenzing","Spark","Spanner","Accumulo","Impala")
scale = c(FALSE,rep(TRUE,18))
pri_index = c(TRUE, TRUE, FALSE, rep(TRUE,4), rep(FALSE,2),rep(TRUE,3),FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE)
sec_index = c(TRUE, FALSE, FALSE, rep(TRUE,3), rep(FALSE,3), TRUE, FALSE, TRUE, FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE)
transactions = c(TRUE, FALSE, FALSE, rep("record",3),rep(FALSE,3),rep("EC, record",3), FALSE, "entity group",rep(FALSE,2),TRUE, "record", FALSE)
joins_analytics = c(TRUE, FALSE, TRUE, "MR", "MR", rep(FALSE,2), rep(TRUE,2), rep(FALSE,2), "MR", rep(FALSE,3),TRUE, "?","MR",TRUE )
integr_constrain = c(TRUE, rep(FALSE,7), rep(TRUE,2),rep(FALSE,2),rep(TRUE,7))
views = c(TRUE, rep(FALSE,2), TRUE, rep(FALSE,5), TRUE, rep(FALSE,4), TRUE, FALSE, TRUE, rep(FALSE,2))
Language = c(TRUE, rep(FALSE,6),rep(TRUE,2),rep(FALSE,3),rep(TRUE,5),FALSE,TRUE)
data_model = c("Tables", rep("key_val",2),"document","ext-records", "document","ext-records",rep("Tables",2),rep("key_val",3),rep("Tables",5),"ext-records","Tables")
type_db = c("sql-like",rep("nosql",1),"batch", rep("nosql",4),rep("sql-like",2),rep("nosql",3), rep("sql-like",1),rep("nosql",1),rep("sql-like",3),rep("nosql",1),"sql-like")

dat1 = data.frame(database = databases, scale = scale, pri_index = pri_index, sec_index = sec_index, transactions = transactions, joins_analytics = joins_analytics, integr_constrain = integr_constrain, views = views, Language = Language, data_model = data_model, type_db = type_db)

library(googleVis)
Table <- gvisTable(dat1,options=list(legend='none',
                                         width = 1000,
                                     height = 430,
                                    fontSize=10, frozenColumns = 1))
Table$html$header <- NULL
Table$html$footer <- NULL
Table$html$jsFooter <- NULL
Table$html$caption <- NULL
print(Table)
```

---

## Relaxing consistency guarantee

<img class="center" src= nosql1.png height=350>

> 1. Ensure high availability
> 2. Support updates

---

## Relaxing consistency guarantee

DataBases: "Everyone MUST see the same thing, either old or new, no matter how long it takes."
NoSQL: "For large applications, we can't afford to wait that long, and maybe it doesn't matter anyway."

New design space: High scalability, high availability, eventual consistency.

---


## Eventual Consistency

- Write conflicts will eventually propogate throughout the system.

> - In absence of updates, all replicas converge towards identical copies.
> - What the application sees in the meantime is sensitive to replication mechanisms.
> - Contrast with RDBMS: Immediate consistency, but maybe deadlocks.

---

## CAP Theorem

> - Consistency
    - Do all applications see the same data?
    
> - Availability 
    - Can I interact with the system in the presence of failures?

> - Partitioning
    - If two sections of your system cannot talk to each other, can they make forward progress on their own?
    - If not, you sacrifice availability
    - If so, you might have to sacrifice consistency

> - Conventional databases assume no partitioning
> - NoSQL systems may sacrfice consistency.

---

## CAP Triangle

<img class="center" src= cap.png height=450>

---

## Features of NoSQL

> 1. Ability to horzontally scale 
> 2. The ability to replicate and partition data over multiple servers
> 3. A simple call level interface (in contrast to SQL binding)
> 4. Weaker concurrency model than ACID (Atomicity, Consistency, Isolation, Durability)
> 5. Efficient used of distributed indexes and RAM
> 6. Dynamically add new attributes 

---


## Horizontal Scalability VS Vertical Scalability

### Horizontal Scalability

Distribute both data and load of operations over many servers, with no RAM or disk shared.

### Vertical Scalability

A database system utilizes many cores and/or CPUs that share RAM and disks.

---

## Data Store categories

> - <b>Key-value Stores:</b> Stores with values and index to find them, based on a programmer defined key.

> - <b>Document stores:</b> Stores documents. The documents are indexed and a simple query mechanism is provided.

> - <b>Wide column stores:</b> Store extensible records that can be partitioned vertically and horizontally across nodes. 

> - <b>Relational Databases:</b> These systems store (and index and query) tuples. 

---


## Key-value stores

### Project Voldemort

- Written in Java, Open source, LinkedIn (most contributions).
- Multi-version concurrency control (MVCC) for
updates.
- Replica updates are asynchronous.
- No guarantee of consistent data
- Guarantee of an up-to-date view if you read a majority of
replicas.
- Optimistic locking for consisten multi-record updates
  - In case of conflict with other process, back out.
- Automatic sharding of data.
- Consistent hashing for distribution
  - data hashed to node K is replicated on node K+1...K+n where n is number of extra copies.
- Nodes can be added or removed from cluster; system adopts automatically.
- Voldemort automatically detects and recovers failed nodes.


---

### Riak

- More functionality than typical key-value store.
- Riak objects can be fetched and stored in JSON
format and thus can have multiple fields (like
documents)
- Only support indices on the primary key. 
- Limits the non-primary fields to storage and fetching as
part of a JSON object. 
- Riak lacks the query mechanisms of the document stores; the only
lookup you can do is on primary key.
- Replication of objects and sharding by
hashing on the primary key. 
- Replica values to be temporarily inconsistent. 
- Tunable consistency by specifying how many replicas
must respond for a successful read.
- Same tuning for successful write.
- Different parts of an application can choose different trade-offs
- Uses consistent hashing. 
- No distinguished node to track status of the system.
- Gossip protocol to track alive nodes and know what data they have- Any node may service a client request. 
- Also includes a map/reduce mechanism to split work over all the nodes in a cluster.

---

### Redis

- The client side does the distributed hashing over servers.
- The servers store data in RAM, but data can be copied to disk for
backup or system shutdown.
- System shutdown needed to add more nodes.
- Like Voldemort, it allows lists and sets to be associated with a key.
- Redis does atomic updates by locking.
- Replication is asynchronous.

---

### Scalaris
- Allows key ranges to be assigned to nodes, rather than simply
hashing to nodes. 
    - A query on a range of values does not need to go to every node.
    - Better load balancing depending on key distributions.

- Replication is synchronous. (copies must be updated before the operation is
complete)
    - Data is guaranteed to be consistent.
- Supports transactions with ACID properties on multiple objects.
- Data is stored in memory, but replication and recovery from node
failures provides durability of the updates.
- Scalaris reads and writes must go to a majority of the replicas before an operation completes. 

---

## Summary of key-value store

```{r xtable, results="asis", echo = FALSE}
databases = c("Voldemort", "Riak", "Redis", "Scalaris", "Tokyo Cabinet","Memcached")
basic_operations = c(rep(TRUE,6))
store_data = c(c("RAM/Disk",5),"Both ram and disk")
Transactions = c(FALSE,FALSE,FALSE, TRUE, TRUE, FALSE)
Replication = c(rep("Asynchronous",3), "Synchronous", "Asynchronous", "Synchronous")
Concurrency = c(rep("MVCC",2), rep("locks",4))

Scalability = c(rep(TRUE,6))
dat = data.frame(database = databases, basic_operations = basic_operations, store_data = store_data, Transactions = Transactions, Replication = Replication, Concurrency = Concurrency, Scalability = Scalability)


library(xtable)
tab <- xtable(dat) 
print(tab, type="html")
```

---


## Document stores

### SimpleDB

### CouchDB

### MongoDB

### Terrastore

### 

